Recurrent event analysis in the presence of real-time high frequency data via random subsampling

### 2023-02-03 letters/response-ref-1-v1.tex

{\bf Response to Referee \#2}

I am incredibly grateful for this review.  I am very happy that the referee found that this paper makes a nice contribution to the field of functional data; we, of course, agree. The application was what drove this methodology so we are also happy to hear that the referee found the application interesting. We next address the reviewers specific comments.

\begin{enumerate}
\item {\it There exists a good amount of recent efforts modeling event data with functional data predictor or joint modeling, so a literature review on relevant papers would be useful. See, for example, papers by Dr. Ciprian Crainiceanu from Johns Hopkins, Dr. Sheng Luo from Duke, and Dr. Jiguo Cao from Simon Fraser.}

\vspace{5mm}
XXXXXX.
We added Section XX to describe the related work and connect our methodology to the existing literature.
\vspace{5mm}

\item {\it It might be worth mentioning existing literature on historical functional linear models, e.g., papers by Dr. Damla Senturk from UCLA and collaborators.}

\vspace{5mm}
XXXXXX.
We added Section XX to describe the related work and connect our methodology to the existing literature.
\vspace{5mm}

\item {\it As with any recent history functional linear models, the size of the window ($\Delta$) is an important parameter. So it would be helpful to see some efforts on developing a data-adaptive method for the selection of $\Delta$.}

\vspace{5mm}
XXXXXX
\vspace{5mm}

\item {\it While in Section 3.6 the paper discusses penalized functional regression to reduce overfitting, it is unclear if the paper actually adopts it in simulation studies and application. It is of interest how the smoothing parameter is selected. My concern is that if no penalty is employed, with more than 30 basis functions, overfitting can be severe and much wider confidence bands are often obtained}

\vspace{5mm}
XXXXXX
\vspace{5mm}

\item {\it While in Section 3.6 the paper discusses penalized functional regression to reduce overfitting, it is unclear if the paper actually adopts it in simulation studies and application. It is of interest how the smoothing parameter is selected. My concern is that if no penalty is employed, with more than 30 basis functions, overfitting can be severe and much wider confidence bands are often obtained}

\vspace{5mm}
XXXXXX
\vspace{5mm}

\item {\it In the theoretical section, it would be helpful to understand what intermediate results Assumption 3.4 will lead to.}

\vspace{5mm}
XXXXXX
\vspace{5mm}

\item {\it In the simulation study, it would make sense to me consider a coefficient function for which the magnitude is not only decaying further away from current time and also becomes 0 at $t-\Delta$. In addition, it might be helpful to use relative errors to quantify the accuracy of estimation of coefficient functions.}

\vspace{5mm}
XXXXXXX
\vspace{5mm}

\item {\it Section 2.1 discusses the rationale of modeling functional predictor without noises. Note that the proposed methodology is actually suitable for data with noises. First, the data application seems to suggest noises are indeed present; see Figure 2 (A). Second, I am not sure I fully understand why in the equation after (3), the conditional probability is conditioning on functional predictor observed up to censoring or truncation time. Isn’t it more sensible to condition on functional predictor observed up to t? Then, the conditional probability will not depend on functional predictor beyond time t.}

\vspace{5mm}
The model is under conditional independence $\eta$.
\vspace{5mm}

\item {\it In Section 3.1, the paper considers modeling the mean function separately at event times and non-event times. Is there some reasoning or justification for this?}

\vspace{5mm}
Separate models in FACE paper so we can use the ; however, it makes minimial difference.  So we cite this.
\vspace{5mm}

\item {\it Minor comments: there are typos and inconsistent naming in the paper, e.g., Proposition 3.6 should actually be Lemma 3.6, AI and ACC are used to refer to activity index.}

\vspace{5mm}
We apologize for the typos and inconsistent naming.  This has been resolved.
\vspace{5mm}

### 2023-02-07 write-up/fda-recurrent.tex
updated related work

>>>>

The use of smart devices and other wearables to passively monitor patients has led to a rapid increase in the number of studies with high-frequency sensor streams that can be conceptualized as studies where measurements are functions.
The development of such applications have been accompanied by intense methodological development in regression models where some covariates are functions.  In this article, we devleop a novel inferential approach to functional regression~\citep{James2002,JamesSilverman2005,muller2005,Ramsay2005,kokoszka2017,doi:10.1198/jasa.2009.tm08564,doi:10.1198/016214507000000527}.

Recently, modeling time-to-event data with functional predictors has received a fair bit of attention.  Functional linear Cox regression model (FLCRM) considers the association between a time-to-event and a set of functional~$\{ X(s) \}$ and scalar predictors~$Z \in \mathbb{R}^p$~\citep{https://doi.org/10.1111/biom.12748}. In this setting, the functional and scalar predictors are observed at baseline and the hazard function depends on a non-parametric baseline, a functional linear model~$\int_S X(s) \beta(s)ds$ and a linear model for the baseline scalar predictors~$Z^\top \gamma$.
Since the linear assumption may be too restrictive,~\cite{doi:10.1080/10618600.2020.1853550} consider a more flexible additive functional Cox model for multivariate functional predictors in which the hazard depends on $\int  F \{s, h_{is}[X_i (s)]\} ds$ where $F$ is an unspecified bivariate twice differntiable function.

In our setting, the high frequency sensor process is an outcome and must therefore be jointly modelled with the time-to-event process.  Recently, \cite{Li2022} proposed a functional approach to joint modelling of multivariate longitudinal biomarkers and a time-to-event outcome~\cite{Tsiatis2004,Rizopoulos2010}.  Here, as in traditional joint models, the longitudinal biomarker is observed at only a few observation times (typically $<10$) and is modelled via a functional principal components analysis.   Shared  process  of  to identify the shared progression pattern and outcome-specific progression patterns of the outcomes.

Here a design-based perspective is taken which subsamples non-event times to  avoid some of the computationally demanding aspects of joint models
Subsampling in the context of functional data analysis for massive data has recently been investigated~\cite{https://doi.org/10.48550/arxiv.2104.03446} in which a functional L-optimality criterion is derived.

The use of historical functional linear models....

The main contribution of this paper is to combine advances in historical functional linear models with techniques from design-based inference to circumvent computationally challenging joint models.

The combination leads to a natural penalized generalized linear model that can be estimated with off-the-shelf software.
